Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations

### 2021-06-11 analysis/paper/paper.Rmd
cleaning up paper.Rmd

---
title: "Title Goes Here"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Although well documented through the Columbia and Snake River hydropower system, little is known about movement and survival of Salmon River juvenile Chinook salmon through their migratory corridors above the hydropower system. With increases in water temperatures, predator abundance, and predator distribution, this knowledge is evermore necessary to characterize limiting factors and recovery potential of this imperiled species. To further understanding of predation impacts on juvenile Chinook Salmon in the Salmon River, specifically, the anthropogenically altered Deadwater Reach, we estimated the population of predatory species, calculated the bioenergetic requirement of predator species to maintain size, and estimated the number of juveniles predated upon within said reach. We then used previously established smolt-to-adult return ratios to understand the impact that predation within the Deadwater Reach has on adult returns to the Lemhi River of Idaho, with broader implications for anadromous populations outside of the Lemhi River specifically. 
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

@@

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

@@

<!-- The actual document text starts here: -->

@@

Here is a citation [@Marwick2017]

@@

# Results
*	Total population estimates of Pike Minnow. Chart or table, but likely a chart.
*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
*	Water temperature plot for the 365 average.
*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
*	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
*	Anything else?

<!-- Here's some example analysis code: -->

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion
*	Here we will need to discuss for each result above: 
    *	How do our results compare to other sources within the literature?
    *	What are the broader implications of each result?
*	Management implications and how these results, analysis, modelling framework could be applied elsewhere?

## Conclusion
*	What are the most important results?
    *	Adult returns
    *	Pop estimate of pike minnow
    *	Total predation on juveniles
*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
*	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment). Is there anything in the literature that discusses habitat preferences for pike minnow? If we speed up velocities and add some cover potentially pike minnow predation success will be lowered. Will these fish just move elsewhere? 
*	The end.

@@

<!-- The following line inserts a page break  -->

>>>>

---
title: "Northern Pikeminnow Abundance in Deadwater Slough and Potential Impacts to Local Chinook Salmon Populations"
author:
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: true
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [biomark]
      correspondence: false
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Michael Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Michael W. Ackerman:
      email: Mike.Ackerman@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib
- packages.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Although well documented through the Columbia and Snake River hydropower system, little is known about movement and survival of Salmon River juvenile Chinook salmon through their migratory corridors above the hydropower system. With increases in water temperatures, predator abundance, and predator distribution, this knowledge is evermore necessary to characterize limiting factors and recovery potential of this imperiled species. To further understanding of predation impacts on juvenile Chinook Salmon in the Salmon River, specifically, the anthropogenically altered Deadwater Reach, we estimated the population of predatory species, calculated the bioenergetic requirement of predator species to maintain size, and estimated the number of juveniles predated upon within said reach. We then used previously established smolt-to-adult return ratios to understand the impact that predation within the Deadwater Reach has on adult returns to the Lemhi River of Idaho, with broader implications for anadromous populations outside of the Lemhi River specifically. 
keywords: |
  northern pikeminnow; Chinook salmon; predation; mark-recapture; bioenergetics
highlights: |
  These are the highlights. 
---

@@

@@

@@

@@

# Results

*	Total population estimates of Pike Minnow. Chart or table, but likely a chart.
*	Gastric lavage sampling results. What percent of the diet is chinook salmon from each sampling event. Pie chart? Mike loves pie charts.
*	Water temperature plot for the 365 average.
*	Bioenergetics results. Total mass consumed over a 365 day period. Plot of daily consumption with day of year on x and mass on y.
*	SAR results. Maybe a plot of multiple scenarios depending on gastric lavage results? And pop estimates if they vary in time.
*	Anything else?

# Discussion

*	Here we will need to discuss for each result above: 
    *	How do our results compare to other sources within the literature?
    *	What are the broader implications of each result?
*	Management implications and how these results, analysis, modelling framework could be applied elsewhere?

## Conclusion

*	What are the most important results?
    *	Adult returns
    *	Pop estimate of pike minnow
    *	Total predation on juveniles
*	We have also presented a novel modeling framework for estimating predation on native, critically endangered anadromous species which can be applied to other areas of interest. John day? Others?
*	Also, this is an anthropogenically created dead water area, which could be a candidate for restoration (removing the impoundment). Is there anything in the literature that discusses habitat preferences for pike minnow? If we speed up velocities and add some cover potentially pike minnow predation success will be lowered. Will these fish just move elsewhere? 
*	The end.

@@
